Management induced changes of soil organic carbon on global croplands

### 2020-12-14 SOCbudgetPaper/SOCbudget.Rmd
update paper part

   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons (management related functions), github.com/pik-piam/mrsoil (soil dynamic related functions) and  github.com/pik-piam/mrvalidation (validation data), which are both based on the MADRaT package (github.com/pik-piam/madrat), a framework which aims to improve reproducibility and transparency in data processing. Model results are accessable under (zenode_link). Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.

@@

output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
    md_extensions: +raw_attribute
  md_document:
    variant: markdown_github
  bookdown::word_document2: default
  rticles::copernicus_article: default
  bookdown::pdf_document2: default
\newpage

@@

```{r child = 'appendix.Rmd'}
```
\newpage

>>>>

   We compile calculations as open-source R packages available at github.com/pik-piam/mrcommons \cite{bodirsky_mrcommons_2020} for the management related functions, github.com/pik-piam/mrsoil \cite{karstens_mrcommons_2020} for soil dynamic related functions and  github.com/pik-piam/mrvalidation \cite{bodirsky_mrvalidation_2020} for validation data, which are both based on the MADRaT package at github.com/pik-piam/madrat \cite{dietrich_madrat_2020}, a framework which aims to improve reproducibility and transparency in data processing. Model results are accessable under (zenode_link). Software code for paper and result prepartion can be found under www.github.com/k4rst3ns/historicalsocmanegement.

@@

appendix: |

  \section{Figures and tables in appendices}

  \subsection{Methods}

    ```{=latex}
  \begin{figure}
    \includegraphics[width=13cm]{../ResultNotebooks/Output/Images/climatezones} 
    \caption{Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines \cite{eggleston_ipcc_2006} and has been reimplented by \cite{carre_background_2010}, which is the source of this data. Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange.}
    \label{fig:CLIMzone}
  \end{figure}
  ```

  ```{=latex}
  \begin{table*}[h]
  \caption{Parameterization of harvested organs and their corresponding residues parts as well as allometric coefficients: This table is mainly based on \cite{bodirsky_n2o_2012} together with simple carbon to dry matter assumptions. Allometric coefficients are used as descriped in \cite{eggleston_ipcc_2006} with HI\textsuperscript{prod} being slope\textsubscript{(T)}, HI\textsuperscript{area} intercept\textsubscript{(T)} and RS R\textsubscript{BG-BIO}.
  }
  \begin{tabular}{lllllllllllll}
  \tophline

           &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
  Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
   \middlehline
  tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
  maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
  trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
  rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
  soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
  rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
  groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
  sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
  oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
  puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
  potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
  cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
  sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
  sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
  others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
  foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
  cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
  \bottomhline
           &                 & \multicolumn{8}{l}{
                                             \begin{minipage}[t]{0.50\columnwidth}
                                                        nr/dm -- nitrogen to dry matter ratio\\
                                                        wm/dm -- wet matter to dry matter ratio\\
                                                        c/dm  -- carbon to dry matter ratio\\
                                                        \end{minipage}} &
                                                                          \multicolumn{3}{l}{
                                                                              \begin{minipage}[t]{0.50\columnwidth}
                                                                                          HI\textsuperscript{area} -- harvest index per area\\
                                                                                          HI\textsuperscript{prod} -- harvest index per production\\
                                                                                          RS -- root:shoot ratio \end{minipage}} 
  \end{tabular}
  (\#tab:c2dm)
  \end{table*}
  ```

  <!-- closs in AWMS - Table {#append:TableclossAWMS} -->

  \subsection{Discussion}

    ```{=latex}
  \begin{figure}
    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/reg_comparisonfigure} 
    \caption{Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010. PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model. We use the climate zone specification of the IPCC \cite{eggleston_ipcc_2006}.}
    \label{fig:SOCreg}
  \end{figure}
  ```

    ```{=latex}
  \begin{figure}
    \includegraphics[width=16cm]{../ResultNotebooks/Output/Images/scenario_litter} 
    \caption{Global $\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines (\cite{calvo_buendia_ipcc_2019}), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in \cite{century_model_2000}. \textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally. Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates.}
    \label{fig:SOClitter}
  \end{figure}
  ```

output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
    md_extensions: +raw_attribute
  md_document:
    variant: markdown_github
  bookdown::word_document2: default
  rticles::copernicus_article: default
  bookdown::pdf_document2: default
---
\newpage

@@

### 2020-12-14 SOCbudgetPaper/appendix.Rmd
update paper part

## map on climate zone used for SCF {#append:climatemap}

## carbon 2 dry matter {#append:Tablec2dm}

## closs in AWMS - Table {#append:TableclossAWMS}

## regional SOC stock in GtC from different sources {#append:regcompare}

```{r captionregSOCtable, echo=FALSE}
captionregSOCtable <- "Modelled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. Note that SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010."
```

```{r regSOCtable, out.width = "100%", fig.cap = captionregSOCtable, echo=FALSE, fig.pos="H"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/reg_comparisonfigure.png")
```

with 
Canada, Australia and New Zealand: CAZ; 
China: CHA; 
European Union: EUR; 
India: IND; 
Japan: JPN; 
Latin America: LAM;
Middle East and north Africa: MEA; 
non-EU member states: NEU;
other Asia: OAS; reforming countries: REF; Sub-Saharan Africa:
SSA; United States: USA.

## global SOC debt in GtC for various litter parameterizations {#append:senslitterpnv}

>>>>

\section{Figures and tables in appendices}

\subsection{Methods}

```{r captionCLIMzone, echo=FALSE}
captionCLIMzone <- "Climate zone map adpated from IPCC: The climate zone classification is based on the classification scheme of the IPCC guidelines ([@eggleston_ipcc_2006]) and has been reimplented by [@carre_background_2010], which is the source of this data. Note that the reduced set, used for the comparison of stock change factors is included in the color code with temperate moist in light blue, temperate dry in dark violett, tropical moist in red and tropical dry in orange."
```

```{r CLIMzone, out.width = "16cm", fig.cap = captionCLIMzone, echo=FALSE}
knitr::include_graphics("../ResultNotebooks/Output/Images/climatezones.png")
```

```{=latex}
\begin{table*}[h]
\caption{bla.
}
\begin{tabular}{lllllllllllll}
 \tophline

           &                                 & \multicolumn{3}{l}{Harvested Organs} & \multicolumn{3}{l}{Above-ground Residues} & \multicolumn{2}{l}{Below-ground Residues} & \multicolumn{3}{l}{Allometric coefficients}                \\
Crop code  & Crop Type                       & nr/dm       & wm/dm      & c/dm      & nr/dm         & wm/dm        & c/dm       & nr/dm                & c/dm               & HI\textsuperscript{area} & HI\textsuperscript{prod} & RS   \\
 \middlehline
tece       & Temperate cereals               & 0.0217      & 1.14       & 0.42      & 0.0074        & 1.11         & 0.42       & 0.0098               & 0.38               & 0.58                     & 1.36                     & 0.24 \\
maiz       & Maize                           & 0.016       & 1.14       & 0.42      & 0.0088        & 1.18         & 0.42       & 0.007                & 0.38               & 0.61                     & 1.03                     & 0.22 \\
trce       & Tropical cereals                & 0.0163      & 1.14       & 0.42      & 0.007         & 1.18         & 0.42       & 0.006                & 0.38               & 0.79                     & 1.06                     & 0.22 \\
rice\_pro  & Rice                            & 0.0128      & 1.15       & 0.42      & 0.007         & 1.11         & 0.42       & 0.009                & 0.38               & 2.46                     & 0.95                     & 0.16 \\
soybean    & Soybean                         & 0.0629      & 1.13       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 1.35                     & 0.93                     & 0.19 \\
rapeseed   & Other oil crops (incl rapeseed) & 0.0334      & 1.08       & 0.42      & 0.0081        & 1.11         & 0.42       & 0.0081               & 0.38               & 0                        & 1.86                     & 0.22 \\
groundnut  & Groundnuts                      & 0.0299      & 1.06       & 0.42      & 0.0224        & 1.11         & 0.42       & 0.008                & 0.38               & 1.54                     & 1.07                     & 0.19 \\
sunflower  & Sunflower                       & 0.0216      & 1.08       & 0.42      & 0.008         & 1.11         & 0.42       & 0.008                & 0.38               & 0                        & 1.86                     & 0.22 \\
oilpalm    & Oilpalms                        & 0.0027      & 1.01       & 0.49      & 0.0052        & 1.11         & 0.48       & 0.0053               & 0.47               & 0                        & 1.86                     & 0.24 \\
puls\_pro  & Pulses                          & 0.0421      & 1.1        & 0.42      & 0.0105        & 1.16         & 0.42       & 0.008                & 0.38               & 0.79                     & 0.89                     & 0.19 \\
potato     & Potatoes                        & 0.0144      & 4.55       & 0.42      & 0.0133        & 6.67         & 0.42       & 0.014                & 0.38               & 1.06                     & 0.1                      & 0.2  \\
cassav\_sp & Tropical roots                  & 0.0053      & 2.95       & 0.42      & 0.0101        & 6.67         & 0.42       & 0.014                & 0.38               & 0                        & 0.85                     & 0.2  \\
sugr\_cane & Sugar beet                      & 0.0024      & 3.7        & 0.42      & 0.008         & 3.82         & 0.42       & 0.008                & 0.38               & 0                        & 0.67                     & 0.07 \\
sugr\_beet & Sugar beet                      & 0.0056      & 4.17       & 0.42      & 0.0176        & 5            & 0.42       & 0.014                & 0.38               & 0                        & 0.54                     & 0.2  \\
others     & Fruits, Vegetables, Nuts          & 0.0267      & 5.49       & 0.42      & 0.0081        & 1.88         & 0.42       & 0.007                & 0.38               & 0                        & 0.39                     & 0.22 \\
foddr      & Forage                          & 0.0201      & 4.29       & 0.42      & 0.0192        & 4.1          & 0.42       & 0.0141               & 0.38               & 0                        & 0.28                     & 0.45 \\
cottn\_pro & Cotton seed                     & 0.0365      & 1.09       & 0.42      & 0.0093        & 1.18         & 0.42       & 0.007                & 0.38               & 0                        & 1.48                     & 0.13 \\
\bottomhline
           &                 & \multicolumn{8}{l}{
                                             \begin{minipage}[t]{0.50\columnwidth}
                                                        nr/dm -- nitrogen to dry matter ratio\\
                                                        wm/dm -- wet matter to dry matter ratio\\
                                                        c/dm  -- carbon to dry matter ratio\\
                                                        \end{minipage}} &
                                                                          \multicolumn{3}{l}{
                                                                              \begin{minipage}[t]{0.50\columnwidth}
                                                                                          HI\textsuperscript{area} -- harvest index per area\\
                                                                                          HI\textsuperscript{prod} -- harvest index per production\\
                                                                                          RS -- root:shoot ratio \end{minipage}} 
\end{tabular}
(\#tab:c2dm)
\end{table*}
```

<!-- closs in AWMS - Table {#append:TableclossAWMS} -->

\subsection{Discussion}

```{r captionregSOCtable, echo=FALSE}
captionregSOCtable <- "Modelled as well as data based estimation for climate zone specific SOC stock in GtC for the first 30 cm of soil aggregated over all land area: SoilGrids, GSOC and WISE do not consider changes over time and rely on soil profile data gather over a long period of time, which makes it hard to pinpoint a specific year to these SOC estimations. In this context they will be compared to modelled data (LPJmL4, this study) for the year 2010. PNV denotes the potential natural vegetation state without considering human cropping activities, calculated as reference stock within our model. We use the climate zone specification of the IPCC [eggleston_ipcc_2006]."
```

```{r regSOCtable, out.width = "16cm", fig.cap = captionregSOCtable, echo=FALSE}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/reg_comparisonfigure.png")
```

```{r captionSOClitter, echo=FALSE}
captionSOClitter <- "Global $\\Delta SOC$ for different litter parameterization choices: Whereas perennial grasses (scenario name \textit{LitterPNV-PerennialGrasses}) are given by the IPCC guidelines ([@calvo_buendia_ipcc_2019]), we used CENTURY configuration file for woody biomass parameterization (see tree.100 in [@century_model_2000]). \textit{LitterPNV-CenturyAverage}) forms the baseline parameterization as used within the study and is calculated as the average over all tree compartments and tree types equally. Litter parameterization has the ability to change global SOC stocks and SOC debt, but is robust concerning trends and sequestration rates."
```

```{r SOClitter, out.width = "16cm", fig.cap = captionSOClitter, echo=FALSE}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/scenario_litter.png")
```

### 2020-12-14 SOCbudgetPaper/discussion.Rmd
update paper part

Furthermore, [@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. [@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account. The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decline in SOC debt (less than 1% change).

@@

Modeling management effects at the global scale, however, comes with parametric and structural uncertainties. High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g. the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs. As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities. [@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources. Following this line of argument, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding. However, with an evaluation of management effects (see Sect. \@ref(sec:ipcccompare)), there is no indication for a systematic overrating of residue biomass. Rather, it is also likely that we are still missing carbon inputs e.g. from cover- and inter-cropping to the soil.

@@

Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes, that most of the SOC dynamic happens in the topsoil. In this regard several aspects are over simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies ([@powlson_limited_2014]) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems. Thirdly, emissions from croplands on organic soils (like peat- and wetlands) are likely underestimate substantially.

@@

To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors ([@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019]) which are based on measurement data for croplands (see Table \@ref(tab:SCFtable)). To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig. \@ref(append:climatemap)).

```{r captionSCFtable, echo=FALSE}
captionSCFtable <- "This table shows different estimates of stock change factors $F_{SCF}$, defined as the SOC stock of managed croplands relative to the counterfactual SOC stock under natural vegetation. We compare average values for four IPCC climate zone classifications. IPCC2006 and IPCC2019 default factors (medium input) as well as altered factors for low and high input of organic matter without any other subsystem consideration (see Table 5.5 in [@eggleston_hs_ipcc_2006], [@calvo_buendia_ipcc_2019] are compared to 1990 and 2010 results for historic input levels of this study."
```

```{r SCFtable, out.width = "16cm", fig.cap = captionSCFtable, echo=FALSE}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/TableSCF_comparison.png")
```

Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines ([@calvo_buendia_ipcc_2019]). Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 ([@eggleston_hs_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), leaving our tropical results in between.
Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially. Our results show still especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks. Considering the impact of irrigation ad fertilization on carbon-poor dryland soils, even factors above 1 (see Fig. \@ref(fig:SOCmaps)(c)) might be not surprising.

@@

The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(tab:SOCtable)).

```{r captionSOCtable, echo=FALSE}
captionSOCtable <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time. This makes it hard to pinpoint a specific year for these SOC estimations. In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010."
```

```{r SOCtable, out.width = "8cm", fig.cap = captionSOCtable, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
```

The global estimates of SOC stock from this study are on the lower end compared to other modeled results or data driven estimates. Looking at regional results in Fig. \@ref(append:regcompare), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions. To avoid that this bias influences our results, our study focuses on SOC changes on cropland. Pristine natural vegetated areas without human land management cancel out in calculation of $\Delta SOC$ and $F^\mathrm{SCF}$.

@@

We conducted a sensitivity analysis (Fig. \@ref(append:senslitterpnv) based on various plant parameterizations from the Century model (see Sect. \@ref(sec:scen_litterpnv)). This shows that the general trend of decreasing SOC debt of ~4 GtC is not altered under various estimates for natural SOC stocks.

>>>>

Furthermore, [@sanderman_soil_2017] modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. [@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yields-levels or no-tillage areas were taken into account. The study moreover concludes that crop productivity gains (increasing yield levels by 18%) do not lead to a substantial decline in SOC debt (less than 1\% change).

@@

Modeling management effects at the global scale, however, comes with parametric and structural uncertainties. High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g. the temperate croplands of Central Europe and UK, indicate suspicious high estimations of SOC inputs. As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below and above ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study). [@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources. Following this line of argument, SOC results calculated in this study may overestimate actual SOC stocks due to an overestimation in residue biomass, for which the ratio compared to the harvested organ is normally reduced due to breeding. However, with an evaluation of management effects (see Sect. \@ref(sec:ipcccompare)), there is no indication for a systematic overrating of residue biomass. Rather, it is also likely that we are still missing carbon inputs e.g. from cover- and inter-cropping to the soil.

@@

Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes, that most of the SOC dynamic happens in the topsoil. In this regard several aspects are over simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Secondly changes to the subsoil due to tillage are neglected, which might be important to noticed, hence studies ([@powlson_limited_2014]) have shown, the subsoil to be a game changer in evaluating total SOC losses or gains for no-tillage systems. Thirdly, emissions from croplands on organic soils (like peat- and wetlands) are likely underestimate substantially.

@@

To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]) which are based on measurement data for croplands (see Table \@ref(tab:SCFglo)). To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig. \@ref(fig:CLIMzone)).

```{r SCFglo, echo=FALSE, results = 'asis'}
table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
library(xtable)
print(xtable(table$comparison_table, digit=2, comment=FALSE, 
              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general. For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019. leaving our results in very good agreement with the old default assumptions. Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent. Also modelled $F^{SOC}$ have increased for all climates over time."))
```

Whereas our estimates are too low in the two tropical climate zones, temperate zone default factors are overestimated for the refinement of the IPCC guidelines ([@calvo_buendia_ipcc_2019]). Admittedly, the tropical factor differ substantially between the IPCC guidelines in 2006 ([@eggleston_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), leaving our tropical results in between.
Taking the oversimplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps , whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanish partially. Our results show still especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks. Considering the impact of irrigation ad fertilization on carbon-poor dryland soils, even factors above 1 (see Fig. \@ref(fig:SOCmaps)(c)) might be not surprising.

@@

The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(fig:SOCglo)).

```{r captionSOCglo, echo=FALSE}
captionSOCglo <- "Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time. This makes it hard to pinpoint a specific year for these SOC estimations. In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010."
```

```{r SOCglo, out.width = "8cm", fig.cap = captionSOCglo, echo=FALSE, fig.pos="h"}
# All defaults
knitr::include_graphics("../ResultNotebooks/Output/Images/glo_comparisonfigure.png")
```

The global estimates of SOC stock from this study are on the lower end compared to other modeled results or data driven estimates. Looking at regional results in Fig. \@ref(fig:SOCreg), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal regions. To avoid that this bias influences our results, our study focuses on SOC changes on cropland. Pristine natural vegetated areas without human land management cancel out in calculation of $\Delta SOC$ and $F^\mathrm{SCF}$.

@@

We conducted a sensitivity analysis (Fig. \@ref(fig:SOClitter)) based on various plant parameterizations from the Century model (see Sect. \@ref(sec:scenlitterpnv)). This shows that the general trend of decreasing SOC debt of ~4 GtC is not altered under various estimates for natural SOC stocks.
